---
title: "R Bootcamp: Session 2"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
# Libraries used in this code
# (Add any libraries you might need here)
```

# Vectors

A vector is simply a list of items that are of the same type. To combine the list of items to a vector, use the `c()` function and separate the items by a comma.

```{r}
v1 <- c(3,45,56,732)
v1

is.numeric(v1)
is.integer(v1)
is.double(v1)
```


```{r}
v2 <- c(3L, 5L, 8L, 10L)
is.numeric(v2)
is.integer(v2)
is.double(v2)

v3 <- c("a", "bye", "ciao", "arrived")
is.numeric(v3)
is.integer(v3)
is.double(v3)
is.character(v3)
```

## Sort a Vector
To sort items in a vector alphabetically or numerically, use the `sort()` function:

```{r}
fruits <- c("banana", "apple", "orange", "mango", "lemon")
numbers <- c(13, 3, 5, 7, 20, 2)

sort(fruits)  # Sort a string
sort(numbers) # Sort numbers
```


## Access Vectors
You can access the vector items by referring to its index number inside brackets []. The first item has index 1, the second item has index 2, and so on:

```{r}
fruits <- c("banana", "apple", "orange")

# Access the first item (banana)
fruits[1]

# Access the first and third item (banana and orange)
fruits[c(1, 3)]

# Access all items except for the first item
fruits[c(-1)]

# Change "banana" to "pear"
fruits[1] <- "pear"

# Print fruits
fruits
```
# Matrix

A matrix is a two dimensional data set with columns and rows.

```{r}
# Create a matrix
thismatrix <- matrix(c(1,2,3,4,5,6), nrow = 3, ncol = 2)

# Print the matrix
thismatrix
```

# Arrays

Array can have more than two dimensions.

```{r}
# An array with one dimension with values ranging from 1 to 24
thisarray <- c(1:24)
thisarray

# An array with more than one dimension
multiarray <- array(thisarray, dim = c(4, 3, 2))
multiarray
```










