---
title: "R Bootcamp: Session 1"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
# Libraries used in this code
# (Add any libraries you might need here)
```

# R Variables

Variables are containers for storing data values. 

In R, a variable is created the moment you first assign a value to it. To assign a value to a variable, use the <- sign. To output (or print) the variable value, simply type the variable name.

## Creating and Using Variables

```{r}

# Creating and playing with variables
A <- 10
B <- 5

C <- A + B
C  # Outputs the result of A + B

```

### Performing Operations

```{r}

var.a <- 2.5
var.b <- 4

# Performing some operations between them
result <- var.a / var.b
result  # Outputs the result of var.a / var.b

result.a <- sqrt(var.a)
result.a  # Outputs the square root of var.a

```

### Printing Variables

```{r}
name <- "John Doe"

name  # Auto-prints the value of the name variable

print(name)  # Useful while scripting and looping
```

####Using the print Command in Loops
The print command is particularly useful in loops and scripts to explicitly output values.

```{r}

for (x in 1:10) {
  x # Doesn't result in any output as it is inside a loop
  print(x)  # Prints the current value of x in each iteration
}

```

## Logical Variables and Operators

Logical variables are used to store boolean values (TRUE or FALSE). They are essential in conditional statements and loops.

### Testing the Value of Some Expressions

```{r}
# Testing the value of some expressions (TRUE or FALSE)
4 < 5     # TRUE
10 > 100  # FALSE
4 == 5    # FALSE

# Assigning the result of a logical expression to a variable
result <- 4 < 5
typeof(result)  # Outputs the type of the result (logical)

resultA <- !TRUE  # Negation of TRUE
resultA  # Outputs FALSE
typeof(resultA)  # Outputs the type of resultA (logical)
```

## Logical Operators
Logical operators like AND (&), OR (|), and NOT (!) are used to perform logical operations on boolean values.

```{r}

# This is an OR expression. If one of the variables is TRUE,
# the expression will return TRUE.
result | resultA  # Outputs TRUE (since result is TRUE)
```

# Process Control

## While loop

With the while loop we can execute a set of statements as long as a condition is TRUE:

```{r}
n = 0

while(n < 10){
  print("Testing the while loop ")
  print(n)
  n <- n + 1
}
```

### Break
With the break statement, we can stop the loop even if the while condition is TRUE:

```{r}
i <- 1
while (i < 6) {
  print(i)
  i <- i + 1
  if (i == 4) {
    break
  }
}
```

`Next` is used to skip a iteration without terminating the loop

## For Loop

A for loop is used for iterating over a sequence:

```{r}
fruits <- list("apple", "banana", "cherry")

for (x in fruits) {
  print(x)
}
```

### Nested loops

It is also possible to place a loop inside another loop. This is called a nested loop:

```{r}
adj <- list("red", "yellow", "tasty")

fruits <- list("apple", "banana", "cherry")
for (x in adj) {
  for (y in fruits) {
    print(paste(x, y))
  }
}
```