---
title: "R Bootcamp: Session 4"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
# Install from CRAN
#install.packages('data.table')
require(data.table)
```

# Plot

The `plot()` function is used to draw points (markers) in a diagram.

```{r}
plot(c(1,2,4,8), c(3,5,8,10))
```


## Sequence of points

```{r}
plot(1:10)
```

## Drawing a line

```{r}
plot(1:10, type="l")
```

## Plot Labels

The `plot()` function also accept other parameters, such as `main`, `xlab` and `ylab`.

```{r}
plot(1:10, main="My Graph", xlab="The x-axis", ylab="The y axis")
```

# Pie Chart

A pie chart is a circular graphical view of data.

```{r}
# Create a vector of pies
x <- c(10,20,30,40)

# Display the pie chart
pie(x)
```

# Bar plot

A bar chart uses rectangular bars to visualize data. Bar charts can be displayed horizontally or vertically. The height or length of the bars are proportional to the values they represent.

```{r}
# x-axis values
x <- c("A", "B", "C", "D")

# y-axis values
y <- c(2, 4, 6, 8)

barplot(y, names.arg = x)
```

# Histogram

Histograms can be created using the `hist()` function in R programming language. This function takes in a vector of values for which the histogram is plotted.

```{r}
require(data.table)
mt <- fread("https://raw.githubusercontent.com/selva86/datasets/master/mtcars.csv")
hist(mt$wt)
```

# GGPlot

```{r}
library(ggplot2)
data("midwest", package = "ggplot2")  # load the data
```

Scatter plot

```{r}
ggplot(midwest, aes(x=area, y=poptotal)) + geom_point()
```

```{r}
g <- ggplot(midwest, aes(x=area, y=poptotal)) + geom_point() + geom_smooth(method="lm")  # set se=FALSE to turnoff confidence bands

g1 <- g + coord_cartesian(xlim=c(0,0.1), ylim=c(0, 1000000))  # zooms in

# Add Title and Labels
g1 + labs(title="Area Vs Population", subtitle="From midwest dataset", y="Population", x="Area", caption="Midwest Demographics")

# or

g1 + ggtitle("Area Vs Population", subtitle="From midwest dataset") + xlab("Area") + ylab("Population")
```